### 2021-02-06 chunks/chunk-pbld2.tex
shortening

\begin{Schunk}
% --begin: "pbld2"
\begin{Sinput}
R> uptakeContext <- ctx3(7, 28.3, 46)
R> d <- lcut(CO2, context = list(uptake = uptakeContext))
\end{Sinput}
%
% --end: "pbld2"
\end{Schunk}