### 2021-04-07 chunks/chunk-fcut.logfact.tex
R output commented & prompt disabled

\begin{Schunk}
% --begin: "fcut.logfact"
\begin{Sinput}
logvec <- c(TRUE, FALSE, TRUE, TRUE)
position <- factor(c("worker", "manager", "worker", "accountant"))
fcut(data.frame(employed = logvec, position = position))
\end{Sinput}
\begin{Soutput}
##      employed not.employed position=accountant position=manager
## [1,]        1            0                   0                0
## [2,]        0            1                   0                1
## [3,]        1            0                   0                0
## [4,]        1            0                   1                0
##      position=worker
## [1,]               1
## [2,]               0
## [3,]               1
## [4,]               0
\end{Soutput}
%
% --end: "fcut.logfact"
\end{Schunk}