### 2021-06-09 chunks/chunk-casestudy-assoc-lcut.tex
Added casestudy

\begin{Schunk}
% --begin: "casestudy-assoc-lcut"
\begin{Sinput}
ldata <- lcut(data)
colnames(ldata)
\end{Sinput}
\begin{Soutput}
##  [1] "position=accountant" "position=manager"    "position=worker"    
##  [4] "ex.sm.age"           "si.sm.age"           "ve.sm.age"          
##  [7] "sm.age"              "ml.sm.age"           "ro.sm.age"          
## [10] "qr.sm.age"           "vr.sm.age"           "ty.me.age"          
## [13] "me.age"              "ml.me.age"           "ro.me.age"          
## [16] "qr.me.age"           "vr.me.age"           "ex.bi.age"          
## [19] "si.bi.age"           "ve.bi.age"           "bi.age"             
## [22] "ml.bi.age"           "ro.bi.age"           "qr.bi.age"          
## [25] "vr.bi.age"           "employed"            "not.employed"       
## [28] "ex.sm.salary"        "si.sm.salary"        "ve.sm.salary"       
## [31] "sm.salary"           "ml.sm.salary"        "ro.sm.salary"       
## [34] "qr.sm.salary"        "vr.sm.salary"        "ty.me.salary"       
## [37] "me.salary"           "ml.me.salary"        "ro.me.salary"       
## [40] "qr.me.salary"        "vr.me.salary"        "ex.bi.salary"       
## [43] "si.bi.salary"        "ve.bi.salary"        "bi.salary"          
## [46] "ml.bi.salary"        "ro.bi.salary"        "qr.bi.salary"       
## [49] "vr.bi.salary"
\end{Soutput}
%
% --end: "casestudy-assoc-lcut"
\end{Schunk}