### 2021-06-09 chunks/chunk-casestudy-assoc-reduce.tex

\begin{Schunk}
% --begin: "casestudy-assoc-reduce"
\begin{Sinput}
reduce(ldata, rules=lrules, ratio=1)
\end{Sinput}
\begin{Soutput}
## Empty rule base
\end{Soutput}
%
% --end: "casestudy-assoc-reduce"
\end{Schunk}