### 2021-06-09 chunks/chunk-casestudy-assoc-searchrules.tex

\begin{Schunk}
% --begin: "casestudy-assoc-searchrules"
\begin{Sinput}
lrules <- searchrules(ldata,
                      lhs=grep('^Species=', colnames(ldata), invert=TRUE),
                      rhs=grep('^Species=', colnames(ldata)),
                      minSupport=0.05)
\end{Sinput}
\begin{Soutput}
## Warning in min(rhs): no non-missing arguments to min; returning Inf
\end{Soutput}
\begin{Soutput}
## Warning in max(rhs): no non-missing arguments to max; returning -Inf
\end{Soutput}
\begin{Sinput}
ldf <- as.data.frame(lrules)
head(ldf[order(ldf$support, decreasing=TRUE), ])
\end{Sinput}
\begin{Soutput}
## Error in order(ldf$support, decreasing = TRUE): argument 1 is not a vector
\end{Soutput}
%
% --end: "casestudy-assoc-searchrules"
\end{Schunk}