### 2021-06-09 chunks/chunk-casestudy-assoc-lcut.tex
updated assoc case study

\begin{Schunk}
% --begin: "casestudy-assoc-lcut"
\begin{Sinput}
ldata <- lcut(iris)
colnames(ldata)
\end{Sinput}
\begin{Soutput}
##  [1] "ex.sm.Sepal.Length" "si.sm.Sepal.Length" "ve.sm.Sepal.Length"
##  [4] "sm.Sepal.Length"    "ml.sm.Sepal.Length" "ro.sm.Sepal.Length"
##  [7] "qr.sm.Sepal.Length" "vr.sm.Sepal.Length" "ty.me.Sepal.Length"
## [10] "me.Sepal.Length"    "ml.me.Sepal.Length" "ro.me.Sepal.Length"
## [13] "qr.me.Sepal.Length" "vr.me.Sepal.Length" "ex.bi.Sepal.Length"
## [16] "si.bi.Sepal.Length" "ve.bi.Sepal.Length" "bi.Sepal.Length"   
## [19] "ml.bi.Sepal.Length" "ro.bi.Sepal.Length" "qr.bi.Sepal.Length"
## [22] "vr.bi.Sepal.Length" "ex.sm.Sepal.Width"  "si.sm.Sepal.Width" 
## [25] "ve.sm.Sepal.Width"  "sm.Sepal.Width"     "ml.sm.Sepal.Width" 
## [28] "ro.sm.Sepal.Width"  "qr.sm.Sepal.Width"  "vr.sm.Sepal.Width" 
## [31] "ty.me.Sepal.Width"  "me.Sepal.Width"     "ml.me.Sepal.Width" 
## [34] "ro.me.Sepal.Width"  "qr.me.Sepal.Width"  "vr.me.Sepal.Width" 
## [37] "ex.bi.Sepal.Width"  "si.bi.Sepal.Width"  "ve.bi.Sepal.Width" 
## [40] "bi.Sepal.Width"     "ml.bi.Sepal.Width"  "ro.bi.Sepal.Width" 
## [43] "qr.bi.Sepal.Width"  "vr.bi.Sepal.Width"  "ex.sm.Petal.Length"
## [46] "si.sm.Petal.Length" "ve.sm.Petal.Length" "sm.Petal.Length"   
## [49] "ml.sm.Petal.Length" "ro.sm.Petal.Length" "qr.sm.Petal.Length"
## [52] "vr.sm.Petal.Length" "ty.me.Petal.Length" "me.Petal.Length"   
## [55] "ml.me.Petal.Length" "ro.me.Petal.Length" "qr.me.Petal.Length"
## [58] "vr.me.Petal.Length" "ex.bi.Petal.Length" "si.bi.Petal.Length"
## [61] "ve.bi.Petal.Length" "bi.Petal.Length"    "ml.bi.Petal.Length"
## [64] "ro.bi.Petal.Length" "qr.bi.Petal.Length" "vr.bi.Petal.Length"
## [67] "ex.sm.Petal.Width"  "si.sm.Petal.Width"  "ve.sm.Petal.Width" 
## [70] "sm.Petal.Width"     "ml.sm.Petal.Width"  "ro.sm.Petal.Width" 
## [73] "qr.sm.Petal.Width"  "vr.sm.Petal.Width"  "ty.me.Petal.Width" 
## [76] "me.Petal.Width"     "ml.me.Petal.Width"  "ro.me.Petal.Width" 
## [79] "qr.me.Petal.Width"  "vr.me.Petal.Width"  "ex.bi.Petal.Width" 
## [82] "si.bi.Petal.Width"  "ve.bi.Petal.Width"  "bi.Petal.Width"    
## [85] "ml.bi.Petal.Width"  "ro.bi.Petal.Width"  "qr.bi.Petal.Width" 
## [88] "vr.bi.Petal.Width"  "Species=setosa"     "Species=versicolor"
## [91] "Species=virginica"
\end{Soutput}
%
% --end: "casestudy-assoc-lcut"
\end{Schunk}