### 2021-06-09 chunks/chunk-casestudy-assoc-reduce.tex

\begin{Schunk}
% --begin: "casestudy-assoc-reduce"
\begin{Sinput}
reduce(ldata, rules=lrules, ratio=1)
\end{Sinput}
\begin{Soutput}
##                                                                                      support
## qr.sm.Petal.Width => Species=setosa                                               0.33333334
## ro.me.Sepal.Width & qr.sm.Petal.Length => Species=setosa                          0.26911455
## ve.sm.Sepal.Length => Species=setosa                                              0.06378584
## ro.me.Sepal.Length & ml.me.Petal.Length & ty.me.Petal.Width => Species=versicolor 0.09947090
## ty.me.Petal.Length => Species=versicolor                                          0.07089734
## ro.bi.Petal.Length => Species=virginica                                           0.17762597
## ml.bi.Petal.Width => Species=virginica                                            0.17133440
## vr.sm.Sepal.Width & ro.bi.Petal.Width => Species=virginica                        0.06244373
## qr.bi.Petal.Length & ro.bi.Petal.Width => Species=virginica                       0.16844945
## vr.bi.Sepal.Length & ro.bi.Petal.Width => Species=virginica                       0.12503819
##                                                                                   lhsSupport
## qr.sm.Petal.Width => Species=setosa                                               0.33333334
## ro.me.Sepal.Width & qr.sm.Petal.Length => Species=setosa                          0.26918504
## ve.sm.Sepal.Length => Species=setosa                                              0.06380372
## ro.me.Sepal.Length & ml.me.Petal.Length & ty.me.Petal.Width => Species=versicolor 0.09948596
## ty.me.Petal.Length => Species=versicolor                                          0.07089734
## ro.bi.Petal.Length => Species=virginica                                           0.17762597
## ml.bi.Petal.Width => Species=virginica                                            0.17133440
## vr.sm.Sepal.Width & ro.bi.Petal.Width => Species=virginica                        0.06244373
## qr.bi.Petal.Length & ro.bi.Petal.Width => Species=virginica                       0.16844945
## vr.bi.Sepal.Length & ro.bi.Petal.Width => Species=virginica                       0.12503819
##                                                                                   rhsSupport
## qr.sm.Petal.Width => Species=setosa                                                0.3333333
## ro.me.Sepal.Width & qr.sm.Petal.Length => Species=setosa                           0.3333333
## ve.sm.Sepal.Length => Species=setosa                                               0.3333333
## ro.me.Sepal.Length & ml.me.Petal.Length & ty.me.Petal.Width => Species=versicolor  0.3333333
## ty.me.Petal.Length => Species=versicolor                                           0.3333333
## ro.bi.Petal.Length => Species=virginica                                            0.3333333
## ml.bi.Petal.Width => Species=virginica                                             0.3333333
## vr.sm.Sepal.Width & ro.bi.Petal.Width => Species=virginica                         0.3333333
## qr.bi.Petal.Length & ro.bi.Petal.Width => Species=virginica                        0.3333333
## vr.bi.Sepal.Length & ro.bi.Petal.Width => Species=virginica                        0.3333333
##                                                                                   confidence
## qr.sm.Petal.Width => Species=setosa                                                1.0000000
## ro.me.Sepal.Width & qr.sm.Petal.Length => Species=setosa                           0.9997382
## ve.sm.Sepal.Length => Species=setosa                                               0.9997197
## ro.me.Sepal.Length & ml.me.Petal.Length & ty.me.Petal.Width => Species=versicolor  0.9998486
## ty.me.Petal.Length => Species=versicolor                                           1.0000000
## ro.bi.Petal.Length => Species=virginica                                            1.0000000
## ml.bi.Petal.Width => Species=virginica                                             1.0000000
## vr.sm.Sepal.Width & ro.bi.Petal.Width => Species=virginica                         1.0000000
## qr.bi.Petal.Length & ro.bi.Petal.Width => Species=virginica                        1.0000000
## vr.bi.Sepal.Length & ro.bi.Petal.Width => Species=virginica                        1.0000000
\end{Soutput}
%
% --end: "casestudy-assoc-reduce"
\end{Schunk}