### 2021-06-09 chunks/chunk-casestudy-assoc-searchrules.tex

\begin{Schunk}
% --begin: "casestudy-assoc-searchrules"
\begin{Sinput}
lrules <- searchrules(ldata,
                      lhs=grep('^Species=', colnames(ldata), invert=TRUE),
                      rhs=grep('^Species=', colnames(ldata)),
                      minSupport=0.05)
ldf <- as.data.frame(lrules)
head(ldf[order(ldf$support, decreasing=TRUE), ])
\end{Sinput}
\begin{Soutput}
##                                        support lhsSupport rhsSupport
## qr.sm.Petal.Width => Species=setosa  0.3333333  0.3333333  0.3333333
## ro.sm.Petal.Length => Species=setosa 0.3333333  0.3333333  0.3333333
## ml.sm.Petal.Length => Species=setosa 0.3329695  0.3329695  0.3333333
## ro.sm.Petal.Width => Species=setosa  0.3329422  0.3329422  0.3333333
## ml.sm.Petal.Width => Species=setosa  0.3291277  0.3291277  0.3333333
## sm.Petal.Length => Species=setosa    0.3230208  0.3230208  0.3333333
##                                      confidence
## qr.sm.Petal.Width => Species=setosa           1
## ro.sm.Petal.Length => Species=setosa          1
## ml.sm.Petal.Length => Species=setosa          1
## ro.sm.Petal.Width => Species=setosa           1
## ml.sm.Petal.Width => Species=setosa           1
## sm.Petal.Length => Species=setosa             1
\end{Soutput}
%
% --end: "casestudy-assoc-searchrules"
\end{Schunk}