### 2021-06-09 chunks/chunk-casestudy-assoc-lcut.tex
...

\begin{Schunk}
% --begin: "casestudy-assoc-lcut"
\begin{Sinput}
ldata <- lcut(iris)
\end{Sinput}
%
% --end: "casestudy-assoc-lcut"
\end{Schunk}