### 2021-06-09 chunks/chunk-casestudy-assoc-reduce.tex

\begin{Schunk}
% --begin: "casestudy-assoc-reduce"
\begin{Sinput}
r <- reduce(ldata, rules=lrules, ratio=1)
rownames(as.data.frame(r))
\end{Sinput}
\begin{Soutput}
##  [1] "qr.sm.Petal.Width => Species=setosa"                                              
##  [2] "ro.me.Sepal.Width & qr.sm.Petal.Length => Species=setosa"                         
##  [3] "ve.sm.Sepal.Length => Species=setosa"                                             
##  [4] "ro.me.Sepal.Length & ml.me.Petal.Length & ty.me.Petal.Width => Species=versicolor"
##  [5] "ty.me.Petal.Length => Species=versicolor"                                         
##  [6] "ro.bi.Petal.Length => Species=virginica"                                          
##  [7] "ml.bi.Petal.Width => Species=virginica"                                           
##  [8] "vr.sm.Sepal.Width & ro.bi.Petal.Width => Species=virginica"                       
##  [9] "qr.bi.Petal.Length & ro.bi.Petal.Width => Species=virginica"                      
## [10] "vr.bi.Sepal.Length & ro.bi.Petal.Width => Species=virginica"
\end{Soutput}
%
% --end: "casestudy-assoc-reduce"
\end{Schunk}