### 2021-06-09 chunks/chunk-casestudy-assoc-searchrules.tex

\begin{Schunk}
% --begin: "casestudy-assoc-searchrules"
\begin{Sinput}
lrules <- searchrules(ldata,
                      lhs=grep("^Species=", colnames(ldata), invert=TRUE),
                      rhs=grep("^Species=", colnames(ldata)),
                      minSupport=0.05,
                      minConfidence=0.8)
ldf <- as.data.frame(lrules)
head(ldf[order(ldf$support, decreasing=TRUE), ], n=10)
\end{Sinput}
\begin{Soutput}
##                                                                                support
## ro.sm.Petal.Length => Species=setosa                                         0.3333333
## qr.sm.Petal.Width => Species=setosa                                          0.3333333
## ml.sm.Petal.Length => Species=setosa                                         0.3329695
## ro.sm.Petal.Width => Species=setosa                                          0.3329422
## ml.sm.Petal.Width => Species=setosa                                          0.3291277
## sm.Petal.Length => Species=setosa                                            0.3230208
## sm.Petal.Width => Species=setosa                                             0.3175236
## ve.sm.Petal.Width => Species=setosa                                          0.2749806
## ro.me.Sepal.Width & qr.sm.Petal.Length => Species=setosa                     0.2691146
## ro.me.Sepal.Width & qr.sm.Petal.Length & vr.sm.Petal.Width => Species=setosa 0.2691146
##                                                                              lhsSupport
## ro.sm.Petal.Length => Species=setosa                                          0.3333333
## qr.sm.Petal.Width => Species=setosa                                           0.3333333
## ml.sm.Petal.Length => Species=setosa                                          0.3329695
## ro.sm.Petal.Width => Species=setosa                                           0.3329422
## ml.sm.Petal.Width => Species=setosa                                           0.3291277
## sm.Petal.Length => Species=setosa                                             0.3230208
## sm.Petal.Width => Species=setosa                                              0.3175236
## ve.sm.Petal.Width => Species=setosa                                           0.2749806
## ro.me.Sepal.Width & qr.sm.Petal.Length => Species=setosa                      0.2691850
## ro.me.Sepal.Width & qr.sm.Petal.Length & vr.sm.Petal.Width => Species=setosa  0.2691850
##                                                                              rhsSupport
## ro.sm.Petal.Length => Species=setosa                                          0.3333333
## qr.sm.Petal.Width => Species=setosa                                           0.3333333
## ml.sm.Petal.Length => Species=setosa                                          0.3333333
## ro.sm.Petal.Width => Species=setosa                                           0.3333333
## ml.sm.Petal.Width => Species=setosa                                           0.3333333
## sm.Petal.Length => Species=setosa                                             0.3333333
## sm.Petal.Width => Species=setosa                                              0.3333333
## ve.sm.Petal.Width => Species=setosa                                           0.3333333
## ro.me.Sepal.Width & qr.sm.Petal.Length => Species=setosa                      0.3333333
## ro.me.Sepal.Width & qr.sm.Petal.Length & vr.sm.Petal.Width => Species=setosa  0.3333333
##                                                                              confidence
## ro.sm.Petal.Length => Species=setosa                                          1.0000000
## qr.sm.Petal.Width => Species=setosa                                           1.0000000
## ml.sm.Petal.Length => Species=setosa                                          1.0000000
## ro.sm.Petal.Width => Species=setosa                                           1.0000000
## ml.sm.Petal.Width => Species=setosa                                           1.0000000
## sm.Petal.Length => Species=setosa                                             1.0000000
## sm.Petal.Width => Species=setosa                                              1.0000000
## ve.sm.Petal.Width => Species=setosa                                           1.0000000
## ro.me.Sepal.Width & qr.sm.Petal.Length => Species=setosa                      0.9997382
## ro.me.Sepal.Width & qr.sm.Petal.Length & vr.sm.Petal.Width => Species=setosa  0.9997382
\end{Soutput}
%
% --end: "casestudy-assoc-searchrules"
\end{Schunk}