### 2021-06-10 chunks/chunk-casestudy-assoc-searchrules.tex
Updated casestudy

\begin{Schunk}
% --begin: "casestudy-assoc-searchrules"
\begin{Sinput}
lrules <- searchrules(ldata,
                      lhs=grep("^Species=", colnames(ldata), invert=TRUE),
                      rhs=grep("^Species=", colnames(ldata)),
                      minSupport=0.05,
                      minConfidence=0.8,
                      n=0, # search for all rules
                      maxLength=5)
\end{Sinput}
%
% --end: "casestudy-assoc-searchrules"
\end{Schunk}