### 2021-06-10 chunks/chunk-casestudy-assoc-viewrules.tex
updated casestudy

\begin{Schunk}
% --begin: "casestudy-assoc-viewrules"
\begin{Sinput}
ldf <- as.data.frame(lrules)
ldf <- ldf[order(ldf$confidence, decreasing=TRUE), ]
head(ldf[, c('support', 'confidence')], n=8)
\end{Sinput}
\begin{Soutput}
##                                                         support confidence
## bi.Petal.Length => Species=virginica                 0.08827589  1.0000000
## bi.Petal.Width => Species=virginica                  0.13908872  1.0000000
## sm.Petal.Width => Species=setosa                     0.31752357  1.0000000
## sm.Petal.Length => Species=setosa                    0.32302085  1.0000000
## lm.Sepal.Length & um.Sepal.Width => Species=setosa   0.05976040  1.0000000
## sm.Sepal.Length & me.Sepal.Width => Species=setosa   0.11999249  0.9984963
## lm.Sepal.Width & um.Petal.Width => Species=virginica 0.06735610  0.9973243
## bi.Sepal.Length => Species=virginica                 0.05485138  0.9967164
\end{Soutput}
%
% --end: "casestudy-assoc-viewrules"
\end{Schunk}