### 2021-06-10 chunks/chunk-casestudy-assoc-arules-search.tex
updated arules

\begin{Schunk}
% --begin: "casestudy-assoc-arules-search"
\begin{Sinput}
tdata <- transactions(adata)
rules <- apriori(tdata, 
                 parameter=list(support=0.05,
                                target='rules'),
                 appearance=list(lhs=grep('Species=', itemLabels(tdata), invert=TRUE, value=TRUE),
                                 rhs=grep('Species=', itemLabels(tdata), value=TRUE)))
\end{Sinput}
\begin{Soutput}
## Apriori
## 
## Parameter specification:
##  confidence minval smax arem  aval originalSupport maxtime support minlen
##         0.8    0.1    1 none FALSE            TRUE       5    0.05      1
##  maxlen target  ext
##      10  rules TRUE
## 
## Algorithmic control:
##  filter tree heap memopt load sort verbose
##     0.1 TRUE TRUE  FALSE TRUE    2    TRUE
## 
## Absolute minimum support count: 7 
## 
## set item appearances ...[23 item(s)] done [0.00s].
## set transactions ...[23 item(s), 150 transaction(s)] done [0.00s].
## sorting and recoding items ... [21 item(s)] done [0.00s].
## creating transaction tree ... done [0.00s].
## checking subsets of size 1 2 3 4 5 done [0.00s].
## writing ... [64 rule(s)] done [0.00s].
## creating S4 object  ... done [0.00s].
\end{Soutput}
%
% --end: "casestudy-assoc-arules-search"
\end{Schunk}