### 2021-06-10 chunks/chunk-casestudy-assoc-arules-search.tex
...

\begin{Schunk}
% --begin: "casestudy-assoc-arules-search"
\begin{Sinput}
tdata <- transactions(adata)
arules <- apriori(tdata, 
                  parameter=list(support=0.05,
                                 confidence=0.8,
                                 target='rules'),
                  appearance=list(lhs=grep('Species=', itemLabels(tdata), invert=TRUE, value=TRUE),
                                  rhs=grep('Species=', itemLabels(tdata), value=TRUE)))
\end{Sinput}
%
% --end: "casestudy-assoc-arules-search"
\end{Schunk}